---
title: "Statistiek Exercises Week 2"
author: "Jeroen Knoester"
output: pdf_document
---

# Summary statistics
## Exercise 1:
a) Create a vector *x* with 98 random values between 1 and 30 (using the "sample" function).
```{r}
x <- sample(1:30, 98, replace = T)
x
```
b) Create another vector, *x2*, by adding at the end of *x* two values: 50 and -5.
```{r}
x2 <- c(x, -5, 50)
x2
```
c) Verify that the median of the two vectors is the same. Why does that happen?
```{r}
median(x) == median(x2)
```
The median of the 2 vectors is the same, because -5 is less than 16 and 50 is more than 16. The median in between stays the same.

d) Compute the interquartile ranges of both vectors and use them to find out if there are any
outliers in the two vectors.
```{r}
IQR(x)
IQR(x2)
# Check if there is any number in x that is higher then h or lower then l(the highest and lowest boundry)
h <- mean(x) + IQR(x) * 1.5
l <- mean(x) - IQR(x) * 1.5
x[x > h]
x[x < l]
# Same for x2
h2 <- mean(x2) + IQR(x2) * 1.5
l2 <- mean(x2) - IQR(x2) * 1.5
x2[x2 > h2]
x2[x2 < l2]
```
e) Produce a plot containing the two boxplots (together in the same plot) of *x* and *x2*. Can you
see the outliers (if there were any) identified in step d)?
```{r}
boxplot(x, x2)
# Shows the outliner +50 but not the outliner -5
```

f) Compute the standard deviations of both vectors. Which one is smaller and why?
```{r}
sd(x)
sd(x2)
```

The standard deviation of *x* is smaller then the sd of *x2*. A low standard deviation is closer to 0. Since *x2* has outliners, this has a higher standard deviation.
\newpage
# Visualizations## Exercise 1Import the *eba*1977 dataset (from the ISwR package) into *R*. Study the type and content of its
variables.

a) Compute the mean of *cases* and produce a boxplot. Are there any outliers? What is the
value of the third quartile?
```{r}
library(ISwR)
attach(eba1977)
mean(cases)
boxplot(cases)

casesSorted <- sort(cases)
(casesSorted[18] + casesSorted[19])/2
eba1977[cases > 17,]
eba1977[cases< 1, ]
```
There are no outliers, and the value of the thirt quartile = 11
b) Produce the summary of the whole dataset. How many values (rows of the dataset) for each
age group were included in the dataset? How many values (rows of the dataset) for each
city were included in the dataset?
```{r}
summary(eba1977)
```
Six rows for each city in the dataset.

c) Produce two histograms: one with the number of cases in the city of Fredericia and one
with the city of Vejle. Put the two histograms in the same plot. What can you say by
comparing these two histograms?

```{r}
par(mfrow = c(1:2))
hist(cases[city == "Fredericia"], main = "Fredericia")
hist(cases[city == "Vejle"], main = "Vejle")
```

d) Using just one R command, put in a single plot the four boxplots representing the number of
cases for each of the four cities. Do the same with the age groups.
```{r}
layout(1)
boxplot(cases ~ city)
boxplot(cases~age)
```
e) With just one R command, compute the average of cases for each of the four cities. Do the
same with the ages (i.e., compute the average for each of the age groups). Which is the city
with the highest average? Which is the age group with the highest average?```{r}tapply(cases, city, FUN = mean)tapply(cases, age, FUN = mean)```Fredericia has the highest average cases of the 4 cities, the ages between 65-69 has most average cases