---
title: "Statistiek Exercises Week 2"
author: "Jeroen Knoester"
output: pdf_document
---

# Summary statistics## Exercise 1:a) Create a vector *x* with 98 random values between 1 and 30 (using the "sample" function).
```{r}
x <- sample(1:30, 98, replace = T)
x
```
b) Create another vector, *x2*, by adding at the end of *x* two values: 50 and -5.
```{r}
x2 <- c(x, -5, 50)
x2
```
c) Verify that the median of the two vectors is the same. Why does that happen?
```{r}
median(x) == median(x2)
```
The median of the 2 vectors is the same, because -5 is less than 16 and 50 is more than 16. The median in between stays the same.

d) Compute the interquartile ranges of both vectors and use them to find out if there are any
outliers in the two vectors.
```{r}
IQR(x)
IQR(x2)
# Check if there is any number in x that is higher then h or lower then l(the highest and lowest boundry)
h <- mean(x) + IQR(x) * 1.5
l <- mean(x) - IQR(x) * 1.5
x[x > h]
x[x < l]
# Same for x2
h2 <- mean(x2) + IQR(x2) * 1.5
l2 <- mean(x2) - IQR(x2) * 1.5
x2[x2 > h2]
x2[x2 < l2]
```
e) Produce a plot containing the two boxplots (together in the same plot) of *x* and *x2*. Can you
see the outliers (if there were any) identified in step d)?
```{r}
boxplot(x, x2)
# Shows the outliner +50 but not the outliner -5
```

f) Compute the standard deviations of both vectors. Which one is smaller and why?
```{r}
sd(x)
sd(x2)
```

The standard deviation of *x* is smaller then the sd of *x2*. A low standard deviation is closer to 0. Since *x2* has outliners, this has a higher standard deviation.